---
title: "Big Data Analytics using R - Coursework 1"
author: "JP Sainsbury"
date: "15 October 2017"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. (a) Suppose you ask a group of 10 students at Birkbeck College how many brothers and sisters	they have.	The number obtained are as follows:2 3 0 5 2 1 1 0 3 3 Find the following measures of central	tendency:	 (i) the mean,	(ii) the median	and	(iii) the	mode. 

Sample
```{r sample mean}
n = 10
x = c(2, 3, 0, 5, 2, 1, 1, 0, 3, 3)
```

(i) Mean
```{r}
mean(x)
```
 (ii)Meadian
```{r}
median(x)
```
 (iii)Mode
```{r}
names(sort(-table(x)))[1]
```

## Find the following measures of spread:
(iv) the variance
```{r}
var(x)
```
(v)
```{r}
sd(x)
```
##  (b) Suppose these 10 students	have the following age: 23 25 18 45 30 21 22 19 29 35	    (i) Find the covariance and correlation between the number of	 siblings	 and their age.	   (ii) Is there a positive or negative or no correlation between the two?	   (iii) Is there causation between the two?	   Justify your answers.	

```{r}
y = c(23, 25, 18, 45, 30, 21, 22, 19, 29, 35)
plot(x,y, xlab="Number of brothers and sisters", ylab="Ages", main="Scatter Plot of Number of siblings vesus age")
```

(i) Covariance 

```{r}
cov(x,y)
```


(i) Correlation
```{r}
cor(x,y)
```
(ii)
There is a positive correlation between the two.
(iii)
The older generation are more likely to have more brothers and sisters. The plot clearly shows the positive correlation between the two.

## 2 Getting familiar with R

(a) Load library
```{r}
library(MASS)

```
Number of rows and columns.

The rows represent the cases and the columns represent the variables

```{r}
nrow(Boston)
ncol(Boston)
```
Columns also known as the variables contain 14 atributes.They are:
```{r}
names(Boston)
```

## Some pairwise scatterplots 
(b)
The below scatterplot displays a positive correlation between the number of rooms per dwelling and the median value. The more rooms a property has, positivley influences the prices of a house.

```{r}
plot(Boston$rm , Boston$medv, xlab="Average number of rooms per dwelling", ylab="median value of owner-occupied homes in $1000s", main = "Rooms vs Median Value")

cor(Boston$rm , Boston$medv, )
```

The below scatterplot displays a positive correlation between the property tax rate and Nitrogen oxides concentration. This idicates the higher the nitrogen oxides concentraction the more property tax paid. This leads me to believe house prices and nitrogen oxides positivly influence each other. It is known property prices are higher in the innercitys, demand for houses, population, traffic and polution are also concentrated in the inner citys.

```{r}
plot(Boston$tax , Boston$nox, xlab="Full-value property-tax rate per $10,000", ylab="Nitrogen oxides concentration (parts per 10 million)" )

 cor(Boston$tax , Boston$nox, )
```
(c) Are any of the predictors associated with per capita crime rate? If so, explain the relationship

```{r}
table(cor(Boston$crim,Boston))
```

Lets order the results

```{r}
sort(cor(Boston$crim,Boston),decreasing = TRUE)
```

Lets pull out the strongest relationship

```{r}
sort(cor(Boston$crim,Boston),decreasing = TRUE)[2]
```

Index of accessibility has the strongest relationship with per capita crime rate. The easier the access to a particular area has a positive correlation on crime rate. The easier it is to access an area, the more lickly it is to have criminal activities take place there.

Do
any
of
the
suburbs
of
Boston
appear
to
have
particularly
high
crime
rates?
Tax
rates?
Pupil-?-teacher
ratios?
Comment
on
the
range
of
each
predictor.

```{r}
boxplot(Boston$crim,col=c("green"), names=c("Crime Boxplot"), main="Boxplot for crime rates")
```

```{r}
boxplot(Boston$tax,col=c("green"), names=c("Tax Boxplot"), main="Boxplot for tax rates")
```

```{r}
boxplot(Boston$ptratio,col=c("green"), names=c("Pupil-teacher ratio by town"), main="Boxplot for Pupil-teacher ratio by town")
```

(e) How
many
of
the
suburbs
in
this
data
set
bound
the
Charles
river?

```{r}
table(Boston$chas)
sum(Boston$chas)
```

(f) What is the median pupil-?-teacher ratio among the towns in this data set?

```{r}
> median(Boston$ptratio[Boston$chas=="1"])
```

(g) Which suburb of Boston has lowest median value of owner occupied homes? What are the values of the other predictors for that suburb, and how do those values compare to the overall ranges for those predictors? Comment on your findings.
